---
title: "Mini-Project 2"
author: "Jessica Jacinto and Anna Ballou"
date: "2019-03-24"
output: 
  html_document:
    code_folding: show
---

```{r, message=FALSE, include = FALSE}
library(tidyverse)
library(ggplot2)
library(qsub)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```
First we selected the variables from the house elections table that were relevant for our investigation. We chose to focus on Florida as it is often a swing state in elections. 
```{r}
election_table <- house_elections %>%
  select(party, state, ge_winner, fec_id) %>%
  filter(state == "FL")
print(election_table)
```
Next we selected the relevant variables from the contributions table. We chose to focus on 24A transaction type. The significance of this will be explained in our analysis. 
```{r}
cont_filter <- contributions %>%
  select(transaction_amt, transaction_type, cand_id) %>%
  filter(transaction_type == "24A") %>%
print(cont_filter)
```

```{r}
new_table <- cont_filter %>%
  rename(fec_id = cand_id) %>%
  group_by(fec_id)
print(new_table)
```
Lastly, we joined the two tables by fec_id (to ensure that only values corresponding to each candidate from the edited contributions table would be used). We also chose to only look at candidates with a documented outcome (either win or loss) and included their total transaction amounts and the type of transaction. 
```{r}
finaltable <- election_table %>%
  inner_join(new_table, by = "fec_id") %>%
  filter(ge_winner == "W" | ge_winner == "N") %>%
  group_by(transaction_type, party, ge_winner) %>%
  select(-state) %>%
  tibble::rowid_to_column() %>%
  spread(key = transaction_type, value = transaction_amt) %>%
  rename(funds_against = '24A')

print(finaltable)
```

```{r}
party_plot <- ggplot(data = finaltable,aes(x = party, y = funds_against, color = ge_winner )) +
  geom_point()
print(party_plot)
```
