---
title: "Mini-Project 2"
author: "Jessica Jacinto and Anna Ballou"
date: "2019-03-24"
output: 
  html_document:
    code_folding: show
---
Link to github repository: 

```{r, message=FALSE, warning = FALSE}
library(tidyverse)
library(ggplot2)
library(ggthemes)
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```
##Data Wrangling
First we selected the variables from the house selection table that were relevant to our investigation. 
```{r, warning = FALSE}
elections_table <- house_elections %>%
  select(state, party) %>%
  filter(state == "IL" | state == "IN" | state == "IA" | state == "KS" | state == "MI" | state == "MN" | state == "MO" | state == "NE" | state == "ND" | state == "OH" | state == "SD" | state == "WI")
```
Next we selected the variables from the contributions table that were relevant to our investigation. We filtered out all transactions amounts greater than zero as we only wanted to examine cases were actual transactions were made. 
```{r}
contributions_table <- contributions %>%
  select(state, transaction_amt) %>%
  filter(transaction_amt > 0)
```  
Lastly, we joined the two tables by state and calculated the average transaction amount for each state. We arranged the table from greatest average transaction to least. 
```{r, warning = FALSE}
final_table <- elections_table %>%
  inner_join(contributions_table, by = c("state")) %>%
  group_by(state) %>%
  summarize(avg_transaction = mean(transaction_amt)) %>%
  arrange(desc(avg_transaction))
print(final_table)
```

##Data Graphic
We plotted state (in the midwest) against the average transaction amount for that state. 
```{r, warning = FALSE}
state_graph <- ggplot(data = final_table, aes(x=state, y=avg_transaction)) +
  geom_bar(mapping = NULL, data = NULL, stat = "identity") +
  scale_y_continuous("Average Transaction Amount ($)", expand=c(0,0)) +
  scale_x_discrete("State") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ggtitle("Average Committee Transactions by State in the Midwest", 
          subtitle = "FEC data") +
  theme(panel.background = element_rect(
    fill = "#89BBD4"))

print(state_graph)
```

##Data Explanation
Higher values of avgerage transaction indicates that a state's committees designate greater amounts of money to politics. Including transaction types from spending against a candidate to loans, a higher average indicates that a state assigned more monetary value to its political climate.

##function
Loop over the individual states in the midwest region. 
```{r, warning = FALSE}
#list of individual states 
midwest = c("IL", "IN", "IA", "KS", "MI", "MN", "MO", "NE", "ND", "OH", "SD", "WI")
south = c("DE", "FL", "GA", "NC", "SC", "VA", "WV", "MD", "AL", "KY", "MI", "TN", "AK", "LA", "OK", "TX")
northeast = c("CT", "ME", "MA", "NH", "RI", "VT", "NJ", "NY", "PA")
west = c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AL", "CA", "HI", "OR", "WA")

```

```{r, warning = FALSE, message = FALSE}

state_list <- function(state_arg){
  elections_individual <- house_elections %>%
  select(state, party) %>%
  filter(state == state_arg)

  contributions_individual <- contributions %>%
  select(state, transaction_amt) %>%
  filter(transaction_amt > 0) %>%
  filter(state == state_arg)


  final_individual <- elections_individual %>%
  inner_join(contributions_individual, by = c("state")) %>%
  group_by(state) %>%
  summarize(avg_transaction = mean(transaction_amt)) %>%
  print(final_individual)
  
}

state_list(west)
state_list(northeast)
state_list(midwest)
state_list(south)
``` 





